ROBUST BAYESIAN RELIABILITY FOR COMPLEX SYSTEMS UNDER PRIOR-DATA CONFLICT

### 2017-06-05 asce-paper.tex
final changes in tex for submission

The choice of prior distribution to encode given expert knowledge is often debatable,
and a specific choice of prior is difficult to justify.
A way to deal with this is to employ sensitivity analysis,
i.e., studying the effect of different choices of prior distribution on the quantities of interest
(in our case, the system reliability function, which, in Bayesian terms, is a posterior predictive distribution).
This idea has been explored in systematic sensitivity analysis, or robust Bayesian methods
(for an overview on this approach, see, e.g.,
\citeNP{1994:berger} or \citeNP{2000:rios}). %\citeNP{2005:ruggeri}, \citeNP{2000:bergerinsuaruggeri}

@@

The paper is organized as follows.
In Section~\ref{sec:weibull}, a Bayesian analysis for Weibull component lifetimes
with fixed shape parameter is described,
illustrating the issue of prior-data conflict.
Section~\ref{sec:modforsurpr} then details the use of sets of priors
for the scale parameter $\lambda_k$ of the Weibull distribution,
showing how this highlights the prior-data conflict issue for $\lambda_k$.
Section~\ref{sec:robrel} then develops a method to calculate system reliability bounds
for a running system based on sets of priors for $\lambda_k$. 
%To this end, Bayesian estimation of $\lambda_k$ under right-censored observations is covered in Section~\ref{sec:lambdawithcens},
%followed by a brief introduction to the survival signature (Section~\ref{sec:sysrelwithsurvsign}).
%Section~\ref{sec:postpred} then details how gathered information on component lifetimes
%is transfomed to the necessary form for use within the survival signature framework,
%and Section~\ref{sec:optimize} explains how lower and upper bounds for $\Rsys(t \mid t > \tnow)$
%are obtained by efficiently optimizing over the parameter sets that define the sets of priors.
Section~\ref{sec:elicitation} discusses elicitation,
by giving guidelines on how to choose prior parameter bounds
that reflect expert information on components.
Section~\ref{sec:examples} contains examples illustrating the merits of our method,
by studying the effect of surprisingly early or late component failures,
showing that observations in conflict to prior assumptions
indeed lead to more cautious system reliability predictions.
Section~\ref{sec:concluding} concludes the paper by summarizing results
and indicating avenues for further research.

@@

Note that this is a general problem in Bayesian analysis with canonical conjugate priors.
For such priors, the same update formula \eqref{eq:ig-update} applies,
and so conflict is averaged out, for details see \citeN{Walter2009a} and \citeN[\S 3.1.4 and \S A.1.2]{diss}.
%
%\iffalse
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{fig1}
\caption{Prior and posterior cdf for $\lambda$ given surprising observations;
the conflict between prior assumptions and data is averaged out,
with a more pointed posterior giving a false sense of certainty.}
\label{fig:weibull-pdc}
\end{figure}
%\fi

@@

Despite this issue of ignoring prior-data conflict,
the tractability of the update step
%(the posterior is again a parametric distribution)
is a very attractive feature of the conjugate setting.
As was shown by \citeN{Walter2009a} (see also \citeNP{diss}, \S 3.1),
it is possible to retain tractability and to have a meaningful reaction to prior-data conflict
when using sets of priors generated by varying both $\nz$ and $\yz$.
Then, the magnitude of the set of posteriors,
and with it the precision of posterior probability statements,
will be sensitive to the degree of prior-data conflict,
i.e.\ leading to more cautious probability statements when prior-data conflict occurs.

@@

Consider observing a system until $\tnow$,
where the system has $K$ different types of components,
and for each type $k$ there are $n_k$ components in the system.
Denoting the number of type $k$ components that have failed by $\tnow$ by $e_k$,
there are $n_k - e_k$ components still functioning at $\tnow$.
We denote the corresponding vector of observations by
\begin{linenomath*}
\begin{align}
\vec{t}^k_{e_k;n_k} &= \big( \underbrace{t^k_1, \ldots, t^k_{e_k}}_{e_k \text{failure times}},
                             \underbrace{\tpnow, \ldots, \tpnow}_{n_k-e_k \text{censored obs.}} \big)\,,
\end{align}
\end{linenomath*}
where $t^+$ indicates a right-censored observation.
%
%***change to general right-censoring? So $n_k-e_k \to c_k$, $e_k+c_k=n_k$,
%introduce sets of observed and censored?

@@

The survival signature \cite{2012:survsign} is a summary of the structure function
for systems with $K$ groups of exchangeable components.
Denoted by $\Phi(l_1,\ldots,l_K)$, with $l_k=0,1,\ldots,n_k$ for $k=1,\ldots,K$,
it is defined as the probability for the event that the system functions
given that precisely $l_k$ of its $n_k$ components of type $k$ function, for each $k\in \{1,\ldots,K\}$.
Essentially, this creates a $K$-dimensional partition for the event $\Tsys > t$,
such that $\Rsys(t) = P(\Tsys > t)$ can be calculated using the law of total probability,
%System survival calculated by %(see Risk Analysis paper eq. (7))
\begin{linenomath*}
\begin{align}
P(\Tsys > t) &= \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} P(\Tsys > t \mid C^1_t = l_1,\ldots, C^K_t = l_K)
                                                                                  P\Big( \bigcap_{k=1}^K \{ C^k_t = l_k\} \Big) \nonumber\\
             &= \sum_{l_1=0}^{n_1} \cdots \sum_{l_K=0}^{n_K} \Phi(l_1,\ldots,l_K) P\Big( \bigcap_{k=1}^K \{ C^k_t = l_k\} \Big) \nonumber\\
%\intertext{ and assuming components of different types are independent,}
             &= \sum_{l_1=0}^{n_1} \cdots \sum_{l_K=0}^{n_K} \Phi(l_1,\ldots,l_K) \prod_{k=1}^K P(C^k_t = l_k)\,,
\label{eq:sysrel-survsign}
\end{align}
\end{linenomath*}
where $P(C^k_t = l_k)$ is the (predictive) probability that exactly $l_k$ components of type $k$ function at time $t$,
and the last equality holds as we assume that components of different types are independent.
Equation~\ref{eq:sysrel-survsign} shows that the survival signature allows to calculate the system reliability $P(\Tsys > t)$
for arbitrary component failure time distributions,
as the probabilities $P(C^k_t = l_k)$ can be determined from any failure time distribution.
Note that for coherent systems,
the survival signature $\Phi(l_1,\ldots,l_K)$ is non-decreasing in each $l_k$.

%
%A priori, we have, for $l_k = 0,1,\ldots, n_k$,
%\begin{linenomath*}
%\begin{align}
%P(C^k_t = l_k\mid\nkz,\ykz)
% &= { n_k \choose l_k} \int [F_k(t \mid\lambda_k)]^{n_k-l_k}
%                        [1 - F_k(t \mid\lambda_k)]^{l_k} f_{\lambda_k}(\lambda_k\mid\nkz,\ykz) \dd \lambda_k
%\end{align}
%\end{linenomath*}
%under the assumption that components of the same type are independent given $\lambda_k$.

@@

%(Should one work with the subsystem consisting only of the non-failed $n_k - e_k$ components?)\\

@@

%Seen as function in $t$, we see that $t$ appears $n_k-e_k-l_k+1$ times,
%once in each summand, in the denominator of the fraction powered to the $\nkn+1$,
%and summands whith even $j$ are decreasing in $t$,
%while summands whith odd $j$ are increasing in $t$.
%Nevertheless, in total $P(C^k_t = l_k\mid\nkz,\ykz, \vec{t}^k_{e_k;n_k})$
%must be decreasing in $t$, as the remaining $n_k-e_k$ components continue to age,
%and so the probability that $l_k$ of them remain operational must decrease.

@@

Equations~\eqref{eq:lrsysdef} and \eqref{eq:ursysdef} seem to suggest that
a full $2K$-dimensional box-constraint optimization is necessary,
but this is not the case.
Remember that $\Phi(l_1,\ldots,l_k)$ from \eqref{eq:sysrel-survsign} is non-decreasing in each of its arguments $l_1,\ldots,l_K$,
so if there is stochastic dominance in $F(l_k \mid \nkz,\ykz,\vec{t}^k_{e_k;n_k})$,
then there is, for each component type $k$,
a prior parameter pair in $\PkZ$ that minimizes system reliability, and
a prior parameter pair in $\PkZ$ that maximizes system reliability,
independently of the other component types. 
%such that system reliability is
%minimized by a distribution that gives most weight to low values of $l_k$, and
%maximized by a distribution that gives most weight to high values of $l_k$.
Indeed, stochastic dominance in $F(l_k \mid \nkz,\ykz,\vec{t}^k_{e_k;n_k})$ is provided for $\ykz$.
To see this, note that $\ykz$ gives the mean expected lifetime for type $k$ components.
Thus, higher values for $\ykz$ mean higher expected lifetimes for the components,
which in turn increases the probability that many components survive until time $t$,
and with it, decreases the propability of few or no components surviving,
so in total giving low probability weight for low values of $l_k$,
and high probability weight for high values of $l_k$. 
Therefore, for any fixed value of $\nkz$, 
the lower bound of $F(l_k \mid \nkz,\ykz,\vec{t}^k_{e_k;n_k})$ for all $l_k$ is obtained with $\ykzu$, and
the upper bound of $F(l_k \mid \nkz,\ykz,\vec{t}^k_{e_k;n_k})$ for all $l_k$ is obtained with $\ykzl$.
%
%***need formal first-order stochastic dominance in $\ykz$ result for $P(C^k_t = l_k\mid\nkz,\ykz, \vec{t}^k_{e_k;n_k})$?
%I have it for inverse gamma posterior $f(\lambda_k\mid \nkz,\ykz, \vec{t}^k_{e_k;n_k})$,
%but not for posterior predictive***
%
There is however no corresponding result for $\nkz$,
such that different values of $\nkz$ may minimize (or maximize) $F(l_k \mid \nkz,\ykz,\vec{t}^k_{e_k;n_k})$ at different $l_k$'s.
Therefore, the $\nkz$ values for lower and upper system reliability bounds are obtained by numeric optimization.
%
\iffalse %---------------------------
\begin{linenomath*}
\begin{align} \label{eq:25}
\lefteqn{%
P(\Tsys > t\mid\{\nkz,\ykz, \mbf{t}^k_{e_k;n_k}\}_{k=1}^K)} \hspace*{10ex}\nonumber\\ 
 &= \sum_{l_1=0}^{n_1-e_1} \cdots \sum_{l_K=0}^{n_K-e_K} \Phi(l_1,\ldots,l_K) \prod_{k=1}^K
    P(C^k_t = l_k\mid\nkz,\ykz, \vec{t}^k_{e_k;n_k})
\end{align}
\end{linenomath*}
we cannot simply plug in all the $\ul{P}(C^k_t = l_k\mid \ldots)$ %and $\ol{P}(C^k_t = l_k\mid \ldots)$
to get $\ul{P}(\Tsys > t\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k})$, % and $\ol{P}(\Tsys > t\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k})$,
as the $\ul{P}(C^k_t = l_k\mid \ldots)$ could correspond to different $(\nkz,\ykz)$ for each $l_k$.
(It would give us a lower bound for $\ul{P}(\Tsys > t\mid\ldots)$, however, but it might be very coarse.)
%
While it is clear that the lower bounds for $\ykz$ will lead to the lower system survival function,
the role of $\nkz$ is not so clear.
We therefore resort to numerical optimization of \eqref{eq:25} over $\{\nkz, k=1,\ldots,K\}$
to obtain the lower bound for $P(\Tsys > t\mid\{\nkz,\ykz, \mbf{t}^k_{e_k;n_k}\}_{k=1}^K)$.
\fi %---------------------------

@@

%\pagebreak
%
%
% Now we start the appendices, with the new section name, "Appendix", and a 
%  new counter, "I", "II", etc.
%
\appendix\label{section:references}
%
% Here's the first appendix, the list of references:
%
\bibliography{refs}
%
%\cite{key}   produces citations with full author list and year 
%\citeNP{key} produces citations with full author list and year, but without enclosing parentheses
%\citeA{key}  produces citations with only the full author list
%\citeN{key}  produces citations with the full author list and year, but
%             which can be used as nouns in a sentence; no parentheses appear around
%             the author names, but only around the year
%\citeyear{key}   produces the year information only, within parentheses
%\citeyearNP{key} produces the year information only
%
%
% And now for some pretty impressive notation.  In this example, I have used
%   the tabular environment to line up the columns in ASCE style.
%   Note that this and all appendices (except the references) start with 
%   the \section command
%
\iffalse
\section{Notation}
\emph{The following symbols are used in this paper:}%\par\vspace{0.10in}
\nopagebreak
\par
\begin{tabular}{r  @{\hspace{1em}=\hspace{1em}}  l}
$D$                    & pile diameter (m); \\
$R$                    & distance (m);      and\\
$C_{\mathrm{Oh\;no!}}$ & fudge factor.
\end{tabular}
\fi

\end{document}

>>>>

The choice of prior distribution to encode given expert knowledge is often debatable,
and a specific choice of prior is difficult to justify.
A way to deal with this is to employ sensitivity analysis,
i.e., studying the effect of different choices of prior distribution on the quantities of interest
(in our case, the system reliability function, which, in Bayesian terms, is a posterior predictive distribution).
This idea has been explored in systematic sensitivity analysis, or robust Bayesian methods
(for an overview on this approach, see, e.g.,
\citeNP{1994:berger} or \citeNP{2000:rios}).

@@

The paper is organized as follows.
In Section~\ref{sec:weibull}, a Bayesian analysis for Weibull component lifetimes
with fixed shape parameter is described,
illustrating the issue of prior-data conflict.
Section~\ref{sec:modforsurpr} then details the use of sets of priors
for the scale parameter $\lambda_k$ of the Weibull distribution,
showing how this highlights the prior-data conflict issue for $\lambda_k$.
Section~\ref{sec:robrel} then develops a method to calculate system reliability bounds
for a running system based on sets of priors for $\lambda_k$. 
Section~\ref{sec:elicitation} discusses elicitation,
by giving guidelines on how to choose prior parameter bounds
that reflect expert information on components.
Section~\ref{sec:examples} contains examples illustrating the merits of our method,
by studying the effect of surprisingly early or late component failures,
showing that observations in conflict to prior assumptions
indeed lead to more cautious system reliability predictions.
Section~\ref{sec:concluding} concludes the paper by summarizing results
and indicating avenues for further research.

@@

Note that this is a general problem in Bayesian analysis with canonical conjugate priors.
For such priors, the same update formula \eqref{eq:ig-update} applies,
and so conflict is averaged out, for details see \citeN{Walter2009a} and \citeN[\S 3.1.4 and \S A.1.2]{diss}.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{fig1}
\caption{Prior and posterior cdf for $\lambda$ given surprising observations;
the conflict between prior assumptions and data is averaged out,
with a more pointed posterior giving a false sense of certainty.}
\label{fig:weibull-pdc}
\end{figure}

@@

Despite this issue of ignoring prior-data conflict,
the tractability of the update step
is a very attractive feature of the conjugate setting.
As was shown by \citeN{Walter2009a} (see also \citeNP{diss}, \S 3.1),
it is possible to retain tractability and to have a meaningful reaction to prior-data conflict
when using sets of priors generated by varying both $\nz$ and $\yz$.
Then, the magnitude of the set of posteriors,
and with it the precision of posterior probability statements,
will be sensitive to the degree of prior-data conflict,
i.e.\ leading to more cautious probability statements when prior-data conflict occurs.

@@

Consider observing a system until $\tnow$,
where the system has $K$ different types of components,
and for each type $k$ there are $n_k$ components in the system.
Denoting the number of type $k$ components that have failed by $\tnow$ by $e_k$,
there are $n_k - e_k$ components still functioning at $\tnow$.
We denote the corresponding vector of observations by
\begin{linenomath*}
\begin{align}
\vec{t}^k_{e_k;n_k} &= \big( \underbrace{t^k_1, \ldots, t^k_{e_k}}_{e_k \text{failure times}},
                             \underbrace{\tpnow, \ldots, \tpnow}_{n_k-e_k \text{censored obs.}} \big)\,,
\end{align}
\end{linenomath*}
where $t^+$ indicates a right-censored observation.

@@

The survival signature \cite{2012:survsign} is a summary of the structure function
for systems with $K$ groups of exchangeable components.
Denoted by $\Phi(l_1,\ldots,l_K)$, with $l_k=0,1,\ldots,n_k$ for $k=1,\ldots,K$,
it is defined as the probability for the event that the system functions
given that precisely $l_k$ of its $n_k$ components of type $k$ function, for each $k\in \{1,\ldots,K\}$.
Essentially, this creates a $K$-dimensional partition for the event $\Tsys > t$,
such that $\Rsys(t) = P(\Tsys > t)$ can be calculated using the law of total probability,
\begin{linenomath*}
\begin{align}
P(\Tsys > t) &= \sum_{l_1=0}^{m_1} \cdots \sum_{l_K=0}^{m_K} P(\Tsys > t \mid C^1_t = l_1,\ldots, C^K_t = l_K)
                                                                                  P\Big( \bigcap_{k=1}^K \{ C^k_t = l_k\} \Big) \nonumber\\
             &= \sum_{l_1=0}^{n_1} \cdots \sum_{l_K=0}^{n_K} \Phi(l_1,\ldots,l_K) P\Big( \bigcap_{k=1}^K \{ C^k_t = l_k\} \Big) \nonumber\\
%\intertext{ and assuming components of different types are independent,}
             &= \sum_{l_1=0}^{n_1} \cdots \sum_{l_K=0}^{n_K} \Phi(l_1,\ldots,l_K) \prod_{k=1}^K P(C^k_t = l_k)\,,
\label{eq:sysrel-survsign}
\end{align}
\end{linenomath*}
where $P(C^k_t = l_k)$ is the (predictive) probability that exactly $l_k$ components of type $k$ function at time $t$,
and the last equality holds as we assume that components of different types are independent.
Equation~\ref{eq:sysrel-survsign} shows that the survival signature allows to calculate the system reliability $P(\Tsys > t)$
for arbitrary component failure time distributions,
as the probabilities $P(C^k_t = l_k)$ can be determined from any failure time distribution.
Note that for coherent systems,
the survival signature $\Phi(l_1,\ldots,l_K)$ is non-decreasing in each $l_k$.

@@

@@

@@

Equations~\eqref{eq:lrsysdef} and \eqref{eq:ursysdef} seem to suggest that
a full $2K$-dimensional box-constraint optimization is necessary,
but this is not the case.
Remember that $\Phi(l_1,\ldots,l_k)$ from \eqref{eq:sysrel-survsign} is non-decreasing in each of its arguments $l_1,\ldots,l_K$,
so if there is stochastic dominance in $F(l_k \mid \nkz,\ykz,\vec{t}^k_{e_k;n_k})$,
then there is, for each component type $k$,
a prior parameter pair in $\PkZ$ that minimizes system reliability, and
a prior parameter pair in $\PkZ$ that maximizes system reliability,
independently of the other component types. 
Indeed, stochastic dominance in $F(l_k \mid \nkz,\ykz,\vec{t}^k_{e_k;n_k})$ is provided for $\ykz$.
To see this, note that $\ykz$ gives the mean expected lifetime for type $k$ components.
Thus, higher values for $\ykz$ mean higher expected lifetimes for the components,
which in turn increases the probability that many components survive until time $t$,
and with it, decreases the propability of few or no components surviving,
so in total giving low probability weight for low values of $l_k$,
and high probability weight for high values of $l_k$. 
Therefore, for any fixed value of $\nkz$, 
the lower bound of $F(l_k \mid \nkz,\ykz,\vec{t}^k_{e_k;n_k})$ for all $l_k$ is obtained with $\ykzu$, and
the upper bound of $F(l_k \mid \nkz,\ykz,\vec{t}^k_{e_k;n_k})$ for all $l_k$ is obtained with $\ykzl$.
There is however no corresponding result for $\nkz$,
such that different values of $\nkz$ may minimize (or maximize) $F(l_k \mid \nkz,\ykz,\vec{t}^k_{e_k;n_k})$ at different $l_k$'s.
Therefore, the $\nkz$ values for lower and upper system reliability bounds are obtained by numeric optimization.
%
\iffalse %---------------------------
\begin{linenomath*}
\begin{align} \label{eq:25}
\lefteqn{%
P(\Tsys > t\mid\{\nkz,\ykz, \mbf{t}^k_{e_k;n_k}\}_{k=1}^K)} \hspace*{10ex}\nonumber\\ 
 &= \sum_{l_1=0}^{n_1-e_1} \cdots \sum_{l_K=0}^{n_K-e_K} \Phi(l_1,\ldots,l_K) \prod_{k=1}^K
    P(C^k_t = l_k\mid\nkz,\ykz, \vec{t}^k_{e_k;n_k})
\end{align}
\end{linenomath*}
we cannot simply plug in all the $\ul{P}(C^k_t = l_k\mid \ldots)$ %and $\ol{P}(C^k_t = l_k\mid \ldots)$
to get $\ul{P}(\Tsys > t\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k})$, % and $\ol{P}(\Tsys > t\mid\nkz,\ykz, \mbf{t}^k_{e_k;n_k})$,
as the $\ul{P}(C^k_t = l_k\mid \ldots)$ could correspond to different $(\nkz,\ykz)$ for each $l_k$.
(It would give us a lower bound for $\ul{P}(\Tsys > t\mid\ldots)$, however, but it might be very coarse.)
%
While it is clear that the lower bounds for $\ykz$ will lead to the lower system survival function,
the role of $\nkz$ is not so clear.
We therefore resort to numerical optimization of \eqref{eq:25} over $\{\nkz, k=1,\ldots,K\}$
to obtain the lower bound for $P(\Tsys > t\mid\{\nkz,\ykz, \mbf{t}^k_{e_k;n_k}\}_{k=1}^K)$.
\fi %---------------------------

@@

\pagebreak
%
% Now we start the appendices, with the new section name, "Appendix", and a 
%  new counter, "I", "II", etc.
\appendix
%
%
% Here's the list of references:
%
% \label{section:references}
\bibliography{refs}
%
\end{document}
